A More Complete Algebra for File Synchronization

### 2015-11-09 2015/paper.tex
Described order sets; first part of proof for intersection of sequences

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\czw}{\fscommand{Z}{W}}

@@

@@

>>>>

\newcommand{\cbrk}{\mathtt{break}}
\newcommand{\fscommand}[2]{{#1#2}}
\newcommand{\fsregcommandchar}[1]{\mathtt{#1}}
\newcommand{\fsregcommand}[2]{\fscommand{\fsregcommandchar{#1}}{\fsregcommandchar{#2}}}
\newcommand{\cbb}{\fsregcommand{\empt}{\empt}}
\newcommand{\cbf}{\fsregcommand{\empt}{F}}
\newcommand{\cbd}{\fsregcommand{\empt}{D}}
\newcommand{\cfb}{\fsregcommand{F}{\empt}}
\newcommand{\cff}{\fsregcommand{F}{F}}
\newcommand{\cfd}{\fsregcommand{F}{D}}
\newcommand{\cdb}{\fsregcommand{D}{\empt}}
\newcommand{\cdf}{\fsregcommand{D}{F}}
\newcommand{\cdd}{\fsregcommand{D}{D}}
\newcommand{\cxy}{\fscommand{X}{Y}}
\newcommand{\cyz}{\fscommand{Y}{Z}}
\newcommand{\cxz}{\fscommand{X}{Z}}
\newcommand{\czw}{\fscommand{Z}{W}}
\newcommand{\cqr}{\fscommand{Q}{R}}

@@

\newcommand{\extset}[1]{\wp{#1}}
\newcommand{\orderset}[1]{\vec{\wp}{#1}}

@@

{\bf Going forward we use the algebra derived from the $|\setd|=1$ case.}

\section{Ordering commands}

We often encounter the case where we only have a set of commands without a sepcified order.
Indeed, it is actually the task of the reconciler to determine whether there is an order
in which updates from multiple replicas can all be applied to a filesystem.

By a {\bf minimal sequence or set} of commands we mean a sequence / set in which
there is at most one command on each path.

If $U$ is a minimal set of commands
that we know can be applied to at least one filesystem in some order without breaking it,
then we can split $U$ into a maximum number of disjunct subsets
(components) so that the paths of two commands from different subsets would never be directly related.
If we aim to order the commands in $U$, then we can focus on order the commands
inside the components, as all permutations of the components are allowed:
the commands in them will commute.

Components containing a single command do not pose a problem when determining
possible orders.
(Note that $\cff$ can only appear in a component of one.)

In each component that is larger than one, the patch are directly related
and the commands form construction or destruction chains.
To order the commands,
select the command with the topmost path, so that there would be no command
on its parent path. If the command removes content at this path, the command
needs to follow all other commands in the component, and
the component is a destruction component.
Otherwise the command with the topmost component needs to precede other commands
in the component, and it is a construction component.
Repeat the above process of splitting commands into subsets on the remaining commands
in the component. Recursively, this way we can determine all allowed orders of the commands.

We use $\orderset{U}$ to denote the set of sequences that contain all commands in $U$
and represent all allowed orders. As a specific case, we can regard a sequence of commands $S$
as a set and determine its $\orderset{S}$.
At times by $\orderset{U}$ we mean an arbitrary sequence in the set.

% (O-D)
{\bf Lemma.} If $U$ contains commands on both $p$ and $\pp$, then
there is a $U^*\in\orderset{U}$ where they are next to each other.

{\bf Lemma.} For any minimal sequence $S$ not equivalent to $\cbrk$,
$\orderset{S} = \extset{S}$.

\section{Soundness}

\section{Completeness}

{\bf Lemma.} \((\forall\FS: S\FS = \empt) \Rightarrow S \equiv \cbrk\)

\section{Update detection}

\section{Reconciliation}

If $A$ and $B$ are minimal sequences and there is at least one filesystem neither of them breaks,
we write $A\circ B$. In this case based on the soundness of the algebra
we also know that $A\nequiv\cbrk$ and $B\nequiv\cbrk$.
Also, if there are commands on path $p$ in both $A$ ($\cxy(p, V_Y)$) and $B$ ($\czw(p, V_W)$)
then their input types must match ($X=Z$).

{\bf Lemma.} Let $A$ and $B$ be two minimal sequences so that $A\circ B$.
Then their intersection can also be applied to any filesystem
either can be applied to without breaking it.

Proof.
First we prove the following.
Mark all commands in $A$ that also appear in $B$.
Then $\exists A^* \in \orderset{A}$ in which the marked commands are at the beginning.

We show $A$ can be transformed via equivalences so that this would be true
by induction on the number of non-marked commands preceding the last marked command.
If it is 0, no transformation is needed.
For the induction step, select the last marked command preceded by a non-marked command.
If the two commute, reverse them, thereby reducing the number on relevant non-marked commands.
If they do not commute, they must be a construction or destruction pair as we know
that $A$ is minimal and $A\nequiv\cbrk$.

Let the marked command be $\cxy(p, V_Y)$ and the preceding unmarked command in $A$ be $\czw(q, V_W)$.
We have two cases and we will proceed by indirection.

If $B$ has a command on $q$, let it be $\cqr(q, V_R)$. As $A\circ B$, we know
that $Q=Z$. From 
O-D (TODO ref)
we also know that there is a $B^*\equiv B$ where $\cxy$ and $\cqr$ are next to each other,
and from the construction and destruction laws we know that $\cxy$ determines that like
$\czw(q, V_W)$ in $A$, any command on $q$ must precede it in $B$. As $B$ is minimal, $B^*$
contains the same commands, and so it is the $\cqr$ command that precedes $\cxy$ in $B^*$.
Going back to the construction and destruction laws we see that the output type of the first command
is always the same, and it is either $D$ or $\empt$. Therefore $R=W$ and $\cqr(q, V_R)=\czw(q, V_W)$
as $|\setd|=|\setb|=1$. This is a contradiction as $\czw$ was not marked, but we see it must also be in $B$.

If $B$ has no command on $q$,
